### 2017-03-23 grlp.tex
towards main result

The approximate linear program (ALP) is obtained by making use of LFA in the LP, i.e., by introducing the new variables $r\in \R^k$ and adding the extra constraint $J=\Phi r$ in \eqref{mdplp} with $\Phi \in \R^{n\times k}$ \citep{SchSei85}.

@@

\min_{r\in \R^k}\, &c^\top \Phi r\,\,\,

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Generalized Reduced Linear Programming}\label{sec:grlp}
\todoc{How about the cleaner ``Linear projected approximate linear programming''}
In this section we introduce the computational model used and the optimization problem
that is the main object of our study.
As discussed in the introduction, we are interested in methods that compute a good approximation to the optimal value function.
As noted earlier, at the expense of a modest additional cost, knowing an $O(\epsilon)$ approximation to $J^*$ at a few states suffices to compute actions of an $O(\epsilon)$-optimal policy. We will take a more general view, and we will consider calculating good approximations to $J^*$ with respect to a weighted $1$-norm, where the weights $c$ form a probability distribution over $\S$. Recall that the weighted $1$-norm $\norm{J}_{1,c}$ of a vector $J\in \R^S$ is defined as $\norm{J}_{1,c}  = \sum_s c(s) |J(s)|$. Note that here and in what follows we identify elements of $\R^\S$ (functions, mapping $\S = \{1,\dots,S\}$ to the reals) with elements of $\R^S$ in the obvious way. This allows us to write e.g. $c^\top J$, which denotes $\sum_s c(s) J(s)$.

To introduce the optimization problem we study first recall that 
the optimal value function $J^*$ is the solution of the fixed point equation $TJ^* = J^*$. 
It follows from the definition of $T$ that $J^* = \max_u T_u J^* \ge T_u J^*$ for any $u$, 
where $\ge$ is the componentwise partial ordering of vectors. 
With some abuse of notation, we also introduce $T_a$ to denote $T_u$ where $u(s) = a$ for any $s\in \S$. 
It follows that $J^* \ge T_a J^*$ for any $a\in \A$ and also that $T = \max_a T_a$, where again the maximization is componentwise.
We call a vector $J$ that satisfies $J \ge T_a J$ for any $a\in A$ \emph{superharmonic}. Note that this is a set of linear inequalities. 
By our note on $T$ and $(T_a)_a$, these inequalities can also be written compactly as $J \ge T J$.
\if0
Since $T$ is monotone (i.e., for any $J_1\le J_2$ it holds that $TJ_1 \le T J_2$) and can be seen to be an $\alpha$-contraction with respect to the maximum norm, if $J$ is superharmonic then $J \ge T J \ge T^2 J \ge \dots \ge J^*$, where the last inequality follows since $T$ is an $\alpha$-contraction and hence $T^n J \to J^*$ in the maximum norm, and hence also componentwise.
Thus, $J^*$ is the ``smallest'' superharmonic function. It follows than that 
for any $c\in \R_+^S \doteq [0,\infty)^S$ and any superharmonic $J$, $c^\top J^*\le c^\top J$ and since $J^*$ is also superharmonic, $\min \{ c^\top J\,:\, J \ge T J \} = c^\top J^*$ and if $c$ is positive-valued then the minimizer of $c^\top J$ amongst all superharminoc functions is $J^*$.
\fi
It is not hard to show then that $J^*$ is the smallest superharminoc function (i.e., for any $J$ superharmonic, $J\ge J^*$). It also follows that for any  $c\in \R_{++}^S \doteq (0,\infty)^S$, the unique solution to the linear program $\min\{ c^\top J \,:\, J \ge T J \} =\min \{ c^\top J\,:\, J \ge T_a J, a\in \A \} $ is $J^*$.

Now, let $\phi_1,\ldots,\phi_k : \S \to \R$ be $k$ basis functions. 
\todoc{I think we should use $d$ to denote the number of basis functions. Or at least $n$. Low priority.} 
The \emph{Approximate Linear Program} (ALP) of \citet{SchSei85} 
is obtained by adding the linear constraints $J = \sum_{i=1}^k r_i \phi_i$ to the above linear program. Eliminating $J$ then gives $\min\{ \sum_i r_i c^\top \phi_i \,:\, \sum_i r_i \phi_i \ge g_a + \alpha \sum_i r_i P_a \phi_i, a\in \A, r = (r_i)\in \R^k \}$.
As noted by \citet{SchSei85}, the linear program is feasible as long as $\one$, defined as the vector with all components being identically equal to one, is in the span of $\{\phi_1,\dots,\phi_k\}$. 
\emph{For the purpose of computations, it is assumed that the values $c^\top \phi_i$, $i=1,\dots, k$ and the values $(P_a \phi_i)(s)$ and $g_a(s)$ can be accessed in constant time.} 
This assumption can be relaxed to assuming that one can access $g_a(s)$ and $\phi_i(s)$ for any $(s,a)$ in constant time, as well as to that one can efficiently sample from $c$, from $P_a(s,\cdot)$ for any $(s,a)$ pair, 
but the details of this are the beyond the scope of the present work. As shown by \citet{ALP}, if $\ralp$ denotes the solution to the above ALP then for $\Jalp \doteq \sum_i \ralp(i) \phi_i \doteq \Phi \ralp$ it holds that $\norm{\Jalp - J^*}_{1,c} \le \frac{2 \epsilon}{1-\alpha}$, where $\epsilon = \inf_r \norm{ J^* - \Phi r }_\infty$ is the error of approximating the optimal value with the span of the basis functions $\phi_1,\dots,\phi_k$ and $\norm{J}_\infty = \max_s |J(s)|$ is the maximum norm and $\Phi \in \R^{S\times k}$ is the matrix formed by $(\phi_1,\dots,\phi_k)$. That the error of approximating $J^*$ with $\Jalp$ is $O(\epsilon)$ is significant: The user can focus on finding a good basis, leaving the search for the ``right'' coefficients to a linear program solver.

While the solving the ALP can be significantly cheaper than solving the LP underlying the MDP
and thus it can be advantageous for moderate-scale MDPs,
 \todoc{Give the big-Oh costs!} the number of constraints in the ALP is $SA$, 
 hence the ALP is still infeasible for huge-scale MDPs. 
%While evaluating the objective and checking a single constraint for constraint violations can be done in constant time, 
%since the number of constraints in the ALP is equal to the number of state-action pairs, 
%the ALP is still intractable in general.
To reduce the number of constraints, we consider a relaxation of ALP 
where the constraints are replaced with positive linear 
combinations of them. 
Recalling that the constraints took the form $J \ge g_a + \alpha P_a J$ (with $J = \Phi r$),
choosing $m$ to be target number of constraints, for $1\le i \le m$, the $i$th new constraint is given by
$\sum_a w_{i,a}^\top J \ge \sum_a w_{i,a}^\top(g_a + \alpha P_a J)$, 
where the choice of $m$ and that of the vectors $w_{i,a}\in \R_+^S$ is left to the user.
Note that this results in a linear program with $k$ variables and $m$ constraints, which can be written as
\begin{align}\label{grlp}
\begin{split}
&\underset{r\in \R^k}{\min}\, \, c^\top \Phi r\\
&\text{s.t.}\mb  \,\sum_a W_a^\top \Phi r\geq \sum_a W_a^\top (g_a + \alpha P_a) \Phi r\,,
\end{split}
\end{align}
where $W_a = (w_{1,a},\dots,w_{m,a}) \in \R_+^{S \times m}$. 
\todoc{Do we need that $(w_{i,a}^\top)_i$ defines a probability distribution?}
Note that the $(i,j)$th entry of the $m\times k$ constraint matrix of the resulting LP is
$\sum_a  w_{i,a}^\top  \phi_j - \alpha \sum_a w_{i,a}^\top P_a \phi_j$ and assuming that $(w_{i,a})_{a}$ has $p$ nonzero
elements, this can be calculated in $O( p )$ time, making the total cost of obtaining the constraint matrix to be $O(mkp)$ regardless the value of $S$ and $A$. \todoc{Somewhere discuss that the cost of obtaining a policy will depend on $A$. 
We can also discuss how $A$ could be folded into states.}

We will call the LP in \eqref{grlp} the \emph{linearly relaxed approximate linear program (LRALP)}. \todoc{Replace GRLP everywhere with LRALP.} Any LP obtained using 
any constraint selection/generation process can be represented by choosing an appropriate binary-valued matrix
$W^\top = (W_1^\top,\dots,W_A^\top)\in \R_+^{m\times SA}$. In particular, when the constraints are selected
in a random process as suggested by \citet{CS}, the matrix $W$ would be a random, binary-valued matrix.

Note that the LRALP may be unbounded. 
Unboundedness could be avoided by adding an extra constraint of the form $r\in \N$ to the LRALP,
for a properly chosen polyhedron $\N \subset \R^k$.%
\footnote{
In particular, to obtain their theoretical result, \citet{CS} need the assumption that the set $\N$ 
is bounded and that it contains $\ralp$. In fact, 
the error bound derived by \citeauthor{CS} depends on the \emph{worst} 
error of approximating $J^*$ with $\Phi r$ when
$r$ ranges over $\N$. Hence, if $\N$ is unbounded, their bound is vacuous. In the context of a particular application, 
\citet{CS} demonstrate that $\N$ can be chosen properly to control this term. 
However, no general construction is presented to choose $\N$.}
However, it seems to us that it is downright misleading to think that guaranteeing a bounded solution 
will also lead to reasonable solutions.
Thus we will stick to the above simple form, forcing a discussion of how $W$ should be chosen to get meaningful results.%
\footnote{
The only question is whether there is some value in adding constraints beyond choosing $W$ properly.
Our position is that the set $\N$ would most likely be chosen based on very little and general information;
the useful knowledge is in choosing $W$, not in some general set $\N$. 
Since randomization does not guarantee bounded solutions, \citet{ALP} needs to use $\N$, which incorporates
all the knowledge which makes the LP bounded. Since 
$\N$
or we could incorporate the choice into $W$.
On the occasion when the LP without the extra $r\in \N$ constraint would be unbounded,
the solution found when $r\in \N$ is added would lie on the boundary of $\N$, which
makes one expect the error of the resulting solution to be rather large.
}

Further insight into the choice of $W$ can be gained by
considering the Langrangians of the ALP and LRALP. To write both LP's in a similar form let us introduce $E = (I_{S\times S},\dots,I_{S\times S})^\top$, where $I_{S\times S}$ is the $S\times S$ identity matrix. Further, let $H:\R^S \times \R^{SA}$ be the operator defined by $(HJ)^\top = ( (T_1 J)^\top, \dots, (T_A J)^\top )$. Note that $H$ is a linear operator. Then, the ALP can be written as $\min\{ c^\top \Phi r \,|\, E \Phi r \ge H \Phi r \}$, while LRALP takes the form  
$\min\{ c^\top \Phi r \,|\, W^\top E \Phi r \ge W^\top H \Phi r \}$. 
Hence, their Langrangians are $\mathcal{L}_{\alp}(r,\lambda) = c^\top \Phi r+\lambda^\top (H\Phi r-E\Phi r)$
$\mathcal{L}_{\lralp}(r,q) = c^\top \Phi r+q^\top W^\top (H\Phi r-E\Phi r)$. Thus, we can view $W q$ as a ``linear approximation''
to the dual variable $\lambda \in \R_+^{SA}$. 
This suggests that perhaps $W$ should be chosen such that it approximates well the optimal dual variable.
If $\Phi$ spans $\R^{S}$, the optimal dual variable $\lambda^*$ is known to be the discounted occupancy measure underlying the optimal policy (Theorem 3.18, \cite{Kall}), suggesting that the role of $W$ is very similar to the role of $\Phi$ excepts that the subspace spanned by the columns of $W$ should ideally be close to $\lambda^*$.

We denote an arbitrary solution to GRLP by $\hr$, and the approximate value function by $\hj=\Phi \hr$ and use $\hu$ to denote the greedy policy w.r.t. $\hj$.\\
\if0
\textbf{Linear function approximation in Primal and Dual Variables:} Let us look at the Lagrangian of ALP and GRLP in
\eqref{lag} and \eqref{lag2} respectively, i.e.,
\begin{align}\label{lag}
\tilde{L}(r,\lambda)=c^\top \Phi r+\lambda^\top (T\Phi r-\Phi r), \\ \label{lag2}\hat{L}(r,q)=c^\top \Phi r+q^\top W^\top (T\Phi r-\Phi r).
\end{align}
Thus, when $Wq = \lambda$, i.e., when $W$ is a set of basis functions that allow
a low dimensional linear representation of the dual variables $\lambda$,
the two problems are the same.
%Note that $ Wq\approx \lambda$ in \eqref{lag2} is linear function approximation of the Lagrange multipliers.
Hence, while ALP employs LFA in its objective function (i.e., use of $\Phi r$), GRLP employs linear approximation both in the objective function ($\Phi r$) as well as the constraints (use of $W$).
%Further, $W$ can be interpreted as the feature matrix that approximates the Lagrange multipliers as $\lambda\approx Wq$, where $\lambda \in \R^{nd}, r\in \R^m$.
To get a sense of how $W$ should be chosen, recall that
the optimal Lagrange multipliers are the discounted number of visits to the ``state-action pairs'' under an optimal policy $u^*$, i.e.,
\begin{align}
\lambda^*(s,u^*(s))&=\big(c^\top(I-\alpha P_{u^*})^{-1}\big)(s)\nn\\
&= \big(c^\top(I+\alpha P_{u^*}+\alpha^2 P_{u^*}^2+\ldots)\big)(s),\nn\\
\lambda^*(s,a)&=0, \qquad \text{for all } a \neq u^*(s),\nn
\end{align}
where $P_{u^*}$ is the probability transition matrix under $u^*$ ($P_{u^*}(s,s') = P_{u^*(s)}(s,s')$, $s,s'\in S$) \cite{dolgov}. Even though we might not have the optimal policy $u^*$ in practice, the fact that $\lambda^*$ is a probability distribution and that it is a linear combination of $\{P_{u^*},P^2_{u^*},\ldots\}$ hints at the kind of features that might be useful for constructing the $W$ matrix.
\fi

>>>>

The approximate linear program (ALP) is obtained by making use of LFA in the LP, i.e., by introducing the new variables $r\in \Re^k$ and adding the extra constraint $J=\Phi r$ in \eqref{mdplp} with $\Phi \in \Re^{n\times k}$ \citep{SchSei85}.

@@

\min_{r\in \Re^k}\, &c^\top \Phi r\,\,\,

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Linearly Relaxed ALP}\label{sec:grlp}
In this section we introduce the computational model used and the ``Linearly Relaxed Approximate Linear Program''
a relaxation of the ALP.

As discussed in the introduction, we are interested in methods that compute a good approximation to the optimal value function.
As noted earlier, at the expense of a modest additional cost, knowing an $O(\epsilon)$ approximation to $J^*$ at a few states suffices to compute actions of an $O(\epsilon)$-optimal policy. We will take a more general view, and we will consider calculating good approximations to $J^*$ with respect to a weighted $1$-norm, where the weights $c$ form a probability distribution over $\S$. Recall that the weighted $1$-norm $\norm{J}_{1,c}$ of a vector $J\in \Re^S$ is defined as $\norm{J}_{1,c}  = \sum_s c(s) |J(s)|$. Note that here and in what follows we identify elements of $\Re^\S$ (functions, mapping $\S = \{1,\dots,S\}$ to the reals) with elements of $\Re^S$ in the obvious way. This allows us to write e.g. $c^\top J$, which denotes $\sum_s c(s) J(s)$.

To introduce the optimization problem we study first recall that 
the optimal value function $J^*$ is the solution of the fixed point equation $TJ^* = J^*$. 
It follows from the definition of $T$ that $J^* = \max_u T_u J^* \ge T_u J^*$ for any $u$, 
where $\ge$ is the componentwise partial ordering of vectors. 
With some abuse of notation, we also introduce $T_a$ to denote $T_u$ where $u(s) = a$ for any $s\in \S$. 
It follows that $J^* \ge T_a J^*$ for any $a\in \A$ and also that $T = \max_a T_a$, where again the maximization is componentwise.
We call a vector $J$ that satisfies $J \ge T_a J$ for any $a\in A$ \emph{superharmonic}. Note that this is a set of linear inequalities. 
By our note on $T$ and $(T_a)_a$, these inequalities can also be written compactly as $J \ge T J$.
\if0
Since $T$ is monotone (i.e., for any $J_1\le J_2$ it holds that $TJ_1 \le T J_2$) and can be seen to be an $\alpha$-contraction with respect to the maximum norm, if $J$ is superharmonic then $J \ge T J \ge T^2 J \ge \dots \ge J^*$, where the last inequality follows since $T$ is an $\alpha$-contraction and hence $T^n J \to J^*$ in the maximum norm, and hence also componentwise.
Thus, $J^*$ is the ``smallest'' superharmonic function. It follows than that 
for any $c\in \Re_+^S \doteq [0,\infty)^S$ and any superharmonic $J$, $c^\top J^*\le c^\top J$ and since $J^*$ is also superharmonic, $\min \{ c^\top J\,:\, J \ge T J \} = c^\top J^*$ and if $c$ is positive-valued then the minimizer of $c^\top J$ amongst all superharminoc functions is $J^*$.
\fi
It is not hard to show then that $J^*$ is the smallest superharminoc function (i.e., for any $J$ superharmonic, $J\ge J^*$). It also follows that for any  $c\in \Re_{++}^S \doteq (0,\infty)^S$, the unique solution to the linear program $\min\{ c^\top J \,:\, J \ge T J \} =\min \{ c^\top J\,:\, J \ge T_a J, a\in \A \} $ is $J^*$.

Now, let $\phi_1,\ldots,\phi_k : \S \to \Re$ be $k$ basis functions. 
\todoc{I think we should use $d$ to denote the number of basis functions. Or at least $n$. Low priority.} 
The \emph{Approximate Linear Program} (ALP) of \citet{SchSei85} 
is obtained by adding the linear constraints $J = \sum_{i=1}^k r_i \phi_i$ to the above linear program. Eliminating $J$ then gives $\min\{ \sum_i r_i c^\top \phi_i \,:\, \sum_i r_i \phi_i \ge g_a + \alpha \sum_i r_i P_a \phi_i, a\in \A, r = (r_i)\in \Re^k \}$.
As noted by \citet{SchSei85}, the linear program is feasible as long as $\one$, defined as the vector with all components being identically equal to one, is in the span of $\{\phi_1,\dots,\phi_k\}$. 
\emph{For the purpose of computations, it is assumed that the values $c^\top \phi_i$, $i=1,\dots, k$ and the values $(P_a \phi_i)(s)$ and $g_a(s)$ can be accessed in constant time.} 
This assumption can be relaxed to assuming that one can access $g_a(s)$ and $\phi_i(s)$ for any $(s,a)$ in constant time, as well as to that one can efficiently sample from $c$, from $P_a(s,\cdot)$ for any $(s,a)$ pair, 
but the details of this are the beyond the scope of the present work. As shown by \citet{ALP}, if $\ralp$ denotes the solution to the above ALP then for $\Jalp \doteq \sum_i \ralp(i) \phi_i \doteq \Phi \ralp$ it holds that $\norm{\Jalp - J^*}_{1,c} \le \frac{2 \epsilon}{1-\alpha}$, where $\epsilon = \inf_r \norm{ J^* - \Phi r }_\infty$ is the error of approximating the optimal value with the span of the basis functions $\phi_1,\dots,\phi_k$ and $\norm{J}_\infty = \max_s |J(s)|$ is the maximum norm and $\Phi \in \Re^{S\times k}$ is the matrix formed by $(\phi_1,\dots,\phi_k)$. That the error of approximating $J^*$ with $\Jalp$ is $O(\epsilon)$ is significant: The user can focus on finding a good basis, leaving the search for the ``right'' coefficients to a linear program solver.

While the solving the ALP can be significantly cheaper than solving the LP underlying the MDP
and thus it can be advantageous for moderate-scale MDPs,
 \todoc{Give the big-Oh costs!} the number of constraints in the ALP is $SA$, 
 hence the ALP is still infeasible for huge-scale MDPs. 
%While evaluating the objective and checking a single constraint for constraint violations can be done in constant time, 
%since the number of constraints in the ALP is equal to the number of state-action pairs, 
%the ALP is still intractable in general.
To reduce the number of constraints, we consider a relaxation of ALP 
where the constraints are replaced with positive linear 
combinations of them. 
Recalling that the constraints took the form $J \ge g_a + \alpha P_a J$ (with $J = \Phi r$),
choosing $m$ to be target number of constraints, for $1\le i \le m$, the $i$th new constraint is given by
$\sum_a w_{i,a}^\top J \ge \sum_a w_{i,a}^\top(g_a + \alpha P_a J)$, 
where the choice of $m$ and that of the vectors $w_{i,a}\in \Re_+^S$ is left to the user.
Note that this results in a linear program with $k$ variables and $m$ constraints, which can be written as
\begin{align}\label{grlp}
\begin{split}
&\underset{r\in \Re^k}{\min}\, \, c^\top \Phi r\\
&\text{s.t.}\mb  \,\sum_a W_a^\top \Phi r\geq \sum_a W_a^\top (g_a + \alpha P_a) \Phi r\,,
\end{split}
\end{align}
where $W_a = (w_{1,a},\dots,w_{m,a}) \in \Re_+^{S \times m}$. 
\todoc{Do we need that $(w_{i,a}^\top)_i$ defines a probability distribution?}
Note that the $(i,j)$th entry of the $m\times k$ constraint matrix of the resulting LP is
$\sum_a  w_{i,a}^\top  \phi_j - \alpha \sum_a w_{i,a}^\top P_a \phi_j$ and assuming that $(w_{i,a})_{a}$ has $p$ nonzero
elements, this can be calculated in $O( p )$ time, making the total cost of obtaining the constraint matrix to be $O(mkp)$ regardless the value of $S$ and $A$. \todoc{Somewhere discuss that the cost of obtaining a policy will depend on $A$. 
We can also discuss how $A$ could be folded into states.}

We will call the LP in \eqref{grlp} the \emph{linearly relaxed approximate linear program (LRALP)}. \todoc{Replace GRLP everywhere with LRALP.} Any LP obtained using 
any constraint selection/generation process can be represented by choosing an appropriate binary-valued matrix
$W^\top = (W_1^\top,\dots,W_A^\top)\in \Re_+^{m\times SA}$. In particular, when the constraints are selected
in a random process as suggested by \citet{CS}, the matrix $W$ would be a random, binary-valued matrix.

Note that the LRALP may be unbounded. 
Unboundedness could be avoided by adding an extra constraint of the form $r\in \N$ to the LRALP,
for a properly chosen polyhedron $\N \subset \Re^k$.%
\footnote{
In particular, to obtain their theoretical result, \citet{CS} need the assumption that the set $\N$ 
is bounded and that it contains $\ralp$. In fact, 
the error bound derived by \citeauthor{CS} depends on the \emph{worst} 
error of approximating $J^*$ with $\Phi r$ when
$r$ ranges over $\N$. Hence, if $\N$ is unbounded, their bound is vacuous. In the context of a particular application, 
\citet{CS} demonstrate that $\N$ can be chosen properly to control this term. 
However, no general construction is presented to choose $\N$.}
However, it seems to us that it is downright misleading to think that guaranteeing a bounded solution 
will also lead to reasonable solutions.
Thus we will stick to the above simple form, forcing a discussion of how $W$ should be chosen to get meaningful results.%
\footnote{
The only question is whether there is some value in adding constraints beyond choosing $W$ properly.
Our position is that the set $\N$ would most likely be chosen based on very little and general information;
the useful knowledge is in choosing $W$, not in some general set $\N$. 
Since randomization does not guarantee bounded solutions, \citet{ALP} needs to use $\N$, which incorporates
all the knowledge which makes the LP bounded. Since 
$\N$
or we could incorporate the choice into $W$.
On the occasion when the LP without the extra $r\in \N$ constraint would be unbounded,
the solution found when $r\in \N$ is added would lie on the boundary of $\N$, which
makes one expect the error of the resulting solution to be rather large.
}

Further insight into the choice of $W$ can be gained by
considering the Langrangians of the ALP and LRALP. To write both LP's in a similar form let us introduce $E = (I_{S\times S},\dots,I_{S\times S})^\top$, where $I_{S\times S}$ is the $S\times S$ identity matrix. Further, let $H:\Re^S \to \Re^{SA}$ be the operator defined by $(HJ)^\top = ( (T_1 J)^\top, \dots, (T_A J)^\top )$. Note that $H$ is a linear operator, which we call the \emph{linear Bellman operator}. 
Then, the ALP can be written as $\min\{ c^\top \Phi r \,|\, E \Phi r \ge H \Phi r \}$, while LRALP takes the form  
$\min\{ c^\top \Phi r \,|\, W^\top E \Phi r \ge W^\top H \Phi r \}$. 
Hence, their Langrangians are $\mathcal{L}_{\alp}(r,\lambda) = c^\top \Phi r+\lambda^\top (H\Phi r-E\Phi r)$
$\mathcal{L}_{\lralp}(r,q) = c^\top \Phi r+q^\top W^\top (H\Phi r-E\Phi r)$. Thus, we can view $W q$ as a ``linear approximation''
to the dual variable $\lambda \in \Re_+^{SA}$. 
This suggests that perhaps $W$ should be chosen such that it approximates well the optimal dual variable.
If $\Phi$ spans $\Re^{S}$, the optimal dual variable $\lambda^*$ is known to be the discounted occupancy measure underlying the optimal policy (Theorem 3.18, \cite{Kall17}), suggesting that the role of $W$ is very similar to the role of $\Phi$ excepts that the subspace spanned by the columns of $W$ should ideally be close to $\lambda^*$.

\if0
\textbf{Linear function approximation in Primal and Dual Variables:} Let us look at the Lagrangian of ALP and GRLP in
\eqref{lag} and \eqref{lag2} respectively, i.e.,
\begin{align}\label{lag}
\tilde{L}(r,\lambda)=c^\top \Phi r+\lambda^\top (T\Phi r-\Phi r), \\ \label{lag2}\hat{L}(r,q)=c^\top \Phi r+q^\top W^\top (T\Phi r-\Phi r).
\end{align}
Thus, when $Wq = \lambda$, i.e., when $W$ is a set of basis functions that allow
a low dimensional linear representation of the dual variables $\lambda$,
the two problems are the same.
%Note that $ Wq\approx \lambda$ in \eqref{lag2} is linear function approximation of the Lagrange multipliers.
Hence, while ALP employs LFA in its objective function (i.e., use of $\Phi r$), GRLP employs linear approximation both in the objective function ($\Phi r$) as well as the constraints (use of $W$).
%Further, $W$ can be interpreted as the feature matrix that approximates the Lagrange multipliers as $\lambda\approx Wq$, where $\lambda \in \Re^{nd}, r\in \Re^m$.
To get a sense of how $W$ should be chosen, recall that
the optimal Lagrange multipliers are the discounted number of visits to the ``state-action pairs'' under an optimal policy $u^*$, i.e.,
\begin{align}
\lambda^*(s,u^*(s))&=\big(c^\top(I-\alpha P_{u^*})^{-1}\big)(s)\nn\\
&= \big(c^\top(I+\alpha P_{u^*}+\alpha^2 P_{u^*}^2+\ldots)\big)(s),\nn\\
\lambda^*(s,a)&=0, \qquad \text{for all } a \neq u^*(s),\nn
\end{align}
where $P_{u^*}$ is the probability transition matrix under $u^*$ ($P_{u^*}(s,s') = P_{u^*(s)}(s,s')$, $s,s'\in S$) \cite{dolgov}. Even though we might not have the optimal policy $u^*$ in practice, the fact that $\lambda^*$ is a probability distribution and that it is a linear combination of $\{P_{u^*},P^2_{u^*},\ldots\}$ hints at the kind of features that might be useful for constructing the $W$ matrix.
\fi

### 2017-03-23 intro.tex
towards main result

The main objective of \emph{planning} is to compute actions of an optimal policy while interacting with an MDP model. 
In finite state-action MDPs,
assuming access to individual transition probabilities and rewards along transitions, 
various algorithms are available to perform this computation in time and space that scales polynomially with the number of states and actions.
However, in most practical applications, the MDP is compactly represented
and if it is not infinite, the number of states scale \emph{exponentially} with the \emph{size of the representation} of the MDP.
%an effect that is known as \emph{Bellman's curse of dimensionality}.
If planners are allowed to perform some fixed amount of calculations for each state encountered,
it is possible to use sampling to make the per-state calculation-cost 
independent of the size of the state space \cite{rust96:randomization,szepesvari2001,kearns2002sparse}.
Nevertheless, the resulting methods are still quite limited. 
In fact, various hardness results show that computing actions of (near-) optimal policies is intractable in various senses
and in various compactly represented MDPs \cite{BlonTsi:00Complexity}.
Given these negative results, 
it is customary to adopt the \emph{modest goal of efficiently computing actions of a policy that 
is nearly as good as the best policy \todoc{This is too ambitious still. Need to tweak.}
in a given restricted policy class}.
Here, within some restrictions (see below), the policy class can be chosen by the user.
The more flexibility the user is given in this choice, the stronger a planning method is.

@@

The  framework of Markov decision processes (MDPs) is useful to mathematically cast optimal sequential decision making problems arising in science and engineering. At any decision instance, an action is made which yields an immediate reward and the system moves to the next state in a stochastic manner such that the next state depends only on the current state and the action chosen. The set of all states, the state space, is denoted by $S$, and the set of all actions, the action space, is denoted by $A$. Formally, a decision rule is called a policy $u$, ($u\colon S\ra A$) and has an associated value function $J_u$\footnote{Without loss of generality value function $J_u$ can be thought of as a vector in $\R^{|S|}$.}, ($J_u(s)\colon S\ra \R$) which specifies the expected cumulative reward obtained by following the policy $u$ starting from each state.\par

@@

This paper is related to LP based techniques for MDPs with large state spaces. One way to handle the large number of states is to restrict the value function to the sub-space spanned by the columns of an $n\times k$ feature matrix $\Phi$\footnote{This is known as linear function approximation (LFA) where in the value function is approximated by $\Phi \tr$, for some $\tr\in \R^k$. The idea of LFA is not restricted to LP based approach and is also widely used in other approximate dynamic programming methods \cite{dpchapter}, which are not discussed in this paper.}. This sub-space restriction can be accommodated in the LP formulation and results in the approximate linear programming (ALP) formulation \cite{ALP,CS,SALP,ALP-Bor}. ALP computes an approximate value function $\tj=\Phi \tr$ and a sub-optimal policy $\tu$ can be obtained using $\tj$\footnote{It is computationally cheap to obtain such a $\tu$ from $\tj$ or $u^*$ from $J^*$}. The sub-optimal policy can then be used to make the decision and results in a cumulative return given by $J_{\tu}$. The performance of ALP was studied in \cite{ALP} in terms of the quantities $\norm{J^*-\tilde{J}}$ and $\norm{J^*-J_{\tu}}$ ($\norm{\cdot}$ is an appropriate norm) known as prediction error and the control error respectively. Here, prediction error is the error in the approximate value function $\tj$ and the control error is the loss in performance due to the sub-optimal policy $\tu$.\par

@@

Linear function approximation (LFA), i.e., letting $\tilde{J}=\Phi \tr$ where $\Phi \in \R^{|S|\times k}$ is a so-called feature matrix and $\tr*\in \R^k$ is a weight vector (to be computed), is the most widely used method of approximation. Here, dimensionality reduction is achieved by choosing $\Phi$ to have fewer columns in comparison to the number of states ($k<<|S|$), holding the promise of being able to work with MDPs regardless of the number of states.\par

>>>>

The main objective of \emph{planning} is to compute actions of an optimal policy while interacting with an MDP model. 
In finite state-action MDPs,
assuming access to individual transition probabilities and rewards along transitions, 
various algorithms are available to perform this computation in time and space that scales polynomially with the number of states and actions.
However, in most practical applications, the MDP is compactly represented
and if it is not infinite, the number of states scale \emph{exponentially} with the \emph{size of the representation} of the MDP.
%an effect that is known as \emph{Bellman's curse of dimensionality}.
If planners are allowed to perform some fixed amount of calculations for each state encountered,
it is possible to use sampling to make the per-state calculation-cost 
independent of the size of the state space \cite{rust96:randomization,szepesvari2001,kearns2002sparse}.
Nevertheless, the resulting methods are still quite limited. 
In fact, various hardness results show that computing actions of (near-) optimal policies is intractable in various senses
and in various compactly represented MDPs \cite{BlonTsi:00Complexity}.
Given these negative results, 
it is customary to adopt the \emph{modest goal of efficiently computing actions of a policy that 
is nearly as good as a policy chosen by a suitable (computationally unbounded, and well-informed) oracle
from a given restricted policy class}. \todoc{Is this better now? We should probably elaborate on the oracle idea space permitting.}
Here, within some restrictions (see below), the policy class can be chosen by the user.
The more flexibility the user is given in this choice, the stronger a planning method is.

@@

The  framework of Markov decision processes (MDPs) is useful to mathematically cast optimal sequential decision making problems arising in science and engineering. At any decision instance, an action is made which yields an immediate reward and the system moves to the next state in a stochastic manner such that the next state depends only on the current state and the action chosen. The set of all states, the state space, is denoted by $S$, and the set of all actions, the action space, is denoted by $A$. Formally, a decision rule is called a policy $u$, ($u\colon S\ra A$) and has an associated value function $J_u$\footnote{Without loss of generality value function $J_u$ can be thought of as a vector in $\Re^{|S|}$.}, ($J_u(s)\colon S\ra \Re$) which specifies the expected cumulative reward obtained by following the policy $u$ starting from each state.\par

@@

This paper is related to LP based techniques for MDPs with large state spaces. One way to handle the large number of states is to restrict the value function to the sub-space spanned by the columns of an $n\times k$ feature matrix $\Phi$\footnote{This is known as linear function approximation (LFA) where in the value function is approximated by $\Phi \tr$, for some $\tr\in \Re^k$. The idea of LFA is not restricted to LP based approach and is also widely used in other approximate dynamic programming methods \cite{dpchapter}, which are not discussed in this paper.}. This sub-space restriction can be accommodated in the LP formulation and results in the approximate linear programming (ALP) formulation \cite{ALP,CS,SALP,ALP-Bor}. ALP computes an approximate value function $\tj=\Phi \tr$ and a sub-optimal policy $\tu$ can be obtained using $\tj$\footnote{It is computationally cheap to obtain such a $\tu$ from $\tj$ or $u^*$ from $J^*$}. The sub-optimal policy can then be used to make the decision and results in a cumulative return given by $J_{\tu}$. The performance of ALP was studied in \cite{ALP} in terms of the quantities $\norm{J^*-\tilde{J}}$ and $\norm{J^*-J_{\tu}}$ ($\norm{\cdot}$ is an appropriate norm) known as prediction error and the control error respectively. Here, prediction error is the error in the approximate value function $\tj$ and the control error is the loss in performance due to the sub-optimal policy $\tu$.\par

@@

Linear function approximation (LFA), i.e., letting $\tilde{J}=\Phi \tr$ where $\Phi \in \Re^{|S|\times k}$ is a so-called feature matrix and $\tr*\in \Re^k$ is a weight vector (to be computed), is the most widely used method of approximation. Here, dimensionality reduction is achieved by choosing $\Phi$ to have fewer columns in comparison to the number of states ($k<<|S|$), holding the promise of being able to work with MDPs regardless of the number of states.\par
